---
title: "RR_proj1"
author: "Swetha Vuppalanchi"
date: "May 13, 2017"
output: html_document
---

## Read data into R

```{r readata}
setwd("~/Coursera/Data Science/Reproducible Research")
activity_data <- read.csv("activity.csv", header = TRUE)
```

```{r load_packages, include=FALSE}
library(dplyr)
```

## Calculate total number of steps per day
```{r aggregate_data}
activity_data_clean <- na.omit(activity_data)
activity_data_total <- summarize(group_by(activity_data_clean, date), steps = sum(steps))
head(activity_data_total)
```

## Plot histogram and bar plot

```{r plots}

hist(activity_data_total$steps, breaks=20, col="red", xlab = "Total Steps taken each day", main = "Distribution of Total Steps taken each day")
     
```

## Calculate mean and median for total steps per day

```{r calculate_mean_median}
activity_data_means <- mean(activity_data_total$steps)
activity_data_means
activity_data_median <- median(activity_data_total$steps)
activity_data_median
```

## Average number of steps taken across days by interval

```{r timeseries_plot}
avg_steps_per_interval <- summarize(group_by(activity_data_clean, interval), steps = mean(steps))
plot(avg_steps_per_interval$interval, avg_steps_per_interval$steps, type = 'l', col = "green", xlab = "Interval", ylab = "Average steps each day", lwd = 2.5, main = "Average steps across all days by interval")

max_steps_interval <- subset(avg_steps_per_interval, steps == max(steps))
max_steps_interval
```

## Calculate missing values and impute missing values with mean

```{r impute_missing_values}
num_missing_values <- sum(is.na(activity_data))
num_missing_values
imputed_data <- transform(activity_data, steps = ifelse(is.na(activity_data$steps), avg_steps_per_interval$steps[match(activity_data$interval, avg_steps_per_interval$interval)], activity_data$steps))
head(imputed_data)
```

## Plot histogram and calculate mean and median values with new imputed dataset

```{r recalculate_mean_median}
activity_data_total_imp <- summarize(group_by(imputed_data, date), steps = sum(steps))

hist(activity_data_total_imp$steps, col="red", breaks=20, xlab = "Total Steps taken each day", main = "Distribution of Total Steps taken each day")

activity_data_imp_means <- mean(activity_data_total_imp$steps)
activity_data_imp_means
activity_data_imp_median <- median(activity_data_total_imp$steps)
activity_data_imp_median
```

## Calculate the differences between mean and median for the original dataset and imputed dataset

```{r calculate_difference}
mean_diff <- abs(activity_data_imp_means - activity_data_means)
mean_diff

median_diff <- abs(activity_data_imp_median - activity_data_median)
median_diff
```
## Plot average number of steps taken across days per interval for weekdays

```{r avg_steps_weekdays}
imputed_data$date <- as.Date(as.character(imputed_data$date), "%Y-%m-%d")
activity_data_transformed <- transform(imputed_data, weekday_ind = factor(ifelse(weekdays(imputed_data$date) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"), 1, 0)))

activity_weekday <- filter(activity_data_transformed, weekday_ind == 1)
activity_weekend <- filter(activity_data_transformed, weekday_ind == 0)

avg_steps_per_interval_weekday <- summarize(group_by(activity_weekday, interval), steps = mean(steps))
avg_steps_per_interval_weekend <- summarize(group_by(activity_weekend, interval), steps = mean(steps))

par(mfrow = c(1,2))
plot(avg_steps_per_interval_weekday$interval, avg_steps_per_interval_weekday$steps, type = 'l', col = "green", xlab = "Interval", ylab = "Average steps each day", lwd = 2.5, main = "Average steps across weekdays")
plot(avg_steps_per_interval_weekend$interval, avg_steps_per_interval_weekend$steps, type = 'l', col = "red", xlab = "Interval", ylab = "Average steps each day", lwd = 2.5, main = "Average steps across weekends ")
```


